---
title: "Lab 6 Comparing means"
author: "SImeon Alfa"
date: "`r Sys.Date()`"
output: github_document
---

Fill out this report and submit it as your completed assignment.

You can view the R markdown script that created this report here: [README.Rmd](README.Rmd)

## $t$-test

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.


```{r}
library(tidyverse)
fish <- read_csv("chap12q19ElectricFish.csv")
fish_long <-
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>%
  mutate(location = str_remove(location, c("species"))) %>%
  print()
```


### Question A

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

```{r t-test}
fish_ttest <- t.test(species ~ location, data = fish_long)
diff(fish_ttest$estimate)
```


ANSWER
NUll Hypothesis: The tributaries have no effect on the number of species electric fish
Alternate Hypothesis: The tributaries have an effect in the number of electric fish
We failed to reject the null hypothesis as p>0.05 and thus its not significantly different.The tributaries have no effect on the number of species of electric fish.

### Question B

> What is the difference in the mean numbers of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this difference in means?

ANSWER
The difference in the mean numbers of species between areas upstream and downstream of a tributary = -1.83 
95% confidence interval of the  difference in means = between -4.59 and 8.25.


### Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

ANSWER
I made the assumption that the paired differences have a normal distribution in populations.

```{r}
fish_long %>%
  ggplot(aes(x = location, y = tributary)) +
  geom_jitter(aes(color = location),
  shape = 16, alpha = 0.9, width = 0.9)
```

## ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees C/log minute. Rates of heat gain for all crabs are provided in the accompanying data file.

### Question D

> Graph the distribution of body temperatures for each crab type:

```{r}
crab <- read_csv("chap15q27FiddlerCrabFans.csv")
crab %>% 
  filter(!is.na(crabType)) %>% 
  ggplot(aes(x = bodyTemperature)) +
  geom_histogram(
    aes(fill = crabType), 
    bins = 9, 
    alpha = 0.5, 
    position = "identity" 
  ) +
  facet_wrap( ~ crabType, ncol = 1) 

```

### Question E

> Does body temperature vary among crab types? State the null and alternative hypothesis, conduct and ANOVA, and interpret the results.

```{r}
aov_crab <-
  aov(bodyTemperature ~ crabType, data = crab)
aov_crab

```
# ANSWER
Null hypothesis:Temperature does not vary between crab types
Alternative hypothesis:Temperature varies between crab types
The results of the ANOVA show that the F value is greater than one and this means the alternative hypothesis is supported. The p-value is <0.05 which means that the null hypothesis is rejected and there is at least one crab type is different from the others.